Phylogenetics and the human microbiome.

### 2013-06-05 sbreview.tex
cranking away...

The human microbiome is the collection of microbial organisms that live inside of and on the surface of humans.
In this section I will review what is known ...

Archaea has been found in the gut.

Genetics.

The microbiome is dynamic in terms of representation but apparently constant in terms of function.
\cite{hmp2012structure}

The microbiome has remarkable metabolic potential.
\cite{qin2010human}
% To understand the impact of gut microbes on human health and well-being it is crucial to assess their genetic potential. Here we describe the Illumina-based metagenomic sequencing, assembly and characterization of 3.3 million non-redundant microbial genes, derived from 576.7 gigabases of sequence, from faecal samples of 124 European individuals. The gene set, ~150 times larger than the human gene complement, contains an overwhelming majority of the prevalent (more frequent) microbial genes of the cohort and probably includes a large proportion of the prevalent human intestinal microbial genes. The genes are largely shared among individuals of the cohort. Over 99% of the genes are bacterial, indicating that the entire cohort harbours between 1,000 and 1,150 prevalent bacterial species and each individual at least 160 such species, which are also largely shared. We define and describe the minimal gut metagenome and the minimal gut bacterial genome in terms of functions present in all individuals and most bacteria, respectively.

Core microbiome \cite{turnbaugh2008core}.
Human microbiome is spatially organized, as can be seen on skin \cite{grice2009topographical}.
Bacterial community variation in human body habitats across space and time \cite{costello2009bacterial}.
Inter-individual versus intra-individual variation \cite{hmp2012structure}.

16s surveys have shown that antibiotics fundamentally disturb microbial communities, resulting in an effect that lasts years.
\cite{jernberg2007long,dethlefsen2008pervasive,jakobsson2010short,dethlefsen2011incomplete}

Major finding: antibiotics.
\cite{dethlefsen2008pervasive,dethlefsen2011incomplete,jakobsson2010short,jernberg2007long}

Obesity

Microbial ecology: human gut microbes associated with obesity \cite{ley2006microbial}

Amish study \cite{zupancic2012analysis}
"Neither BMI nor any metabolic syndrome trait was associated with a particular gut community. Network analysis identified twenty-two bacterial species and four OTUs that were either positively or inversely correlated with metabolic syndrome traits, suggesting that certain members of the gut microbiota may play a role in these metabolic derangements."

pregancy \cite{koren2012host}

Substantial attention was given to whether microbial ``types'' exist.
A high-profile paper from Per Bork's group had enterotypes: Bacteroides, Prevotella or Ruminococcus.
\cite{arumugam2011enterotypes}
Appears that there is a real tendency to cluster around Bacteroides or Prevotella \cite{wu2011linking}

Although the focus of much of the work has been on bacteria and archaea, other groups of life have attracted more attention lately, such as the fungi.
Viruses too
febrile study \cite{wylie2012sequence}

@@

There has been a shift from interest in community dynamics to fine-scale interactions between host and immune system.

hooper2012interactions,

Trait evolution methods are conspicuously absent.

\cite{wylie2012sequence}
\cite{chen2012associating}

% medical
\cite{clarridge2004}
{{I}mpact of 16{S} r{RNA} gene sequence analysis for identification of bacteria on clinical microbiology and infectious diseases}

\cite{phillips2012microbiome}
Microbiome analysis among bats describes influences of host phylogeny, life history, physiology and geography

@@

\section{Discussion}

What we can expect next.
Interactions with the immune system:
\cite{hooper2012interactions}

Clinical applications, but MALDI-TOF

Field is still not yet mature.
Need to think about study design.

>>>>

The human microbiome is the collection of microbial organisms that live inside of and on the surface of humans.
In this section I will review what is known about them and their effects.
More bacteria than human cells.

The microbiome has remarkable metabolic potential.
\cite{qin2010human}
% To understand the impact of gut microbes on human health and well-being it is crucial to assess their genetic potential. Here we describe the Illumina-based metagenomic sequencing, assembly and characterization of 3.3 million non-redundant microbial genes, derived from 576.7 gigabases of sequence, from faecal samples of 124 European individuals. The gene set, ~150 times larger than the human gene complement, contains an overwhelming majority of the prevalent (more frequent) microbial genes of the cohort and probably includes a large proportion of the prevalent human intestinal microbial genes. The genes are largely shared among individuals of the cohort. Over 99% of the genes are bacterial, indicating that the entire cohort harbours between 1,000 and 1,150 prevalent bacterial species and each individual at least 160 such species, which are also largely shared. We define and describe the minimal gut metagenome and the minimal gut bacterial genome in terms of functions present in all individuals and most bacteria, respectively.
The microbiome is dynamic in terms of representation but apparently constant in terms of function.
\cite{hmp2012structure}
Core microbiome \cite{turnbaugh2008core}.

Microbiome is an ecosystem.
It is remarkably diverse, though less diverse than soil.
Human microbiome is spatially organized, as can be seen on skin \cite{grice2009topographical}.
Bacterial community variation in human body habitats across space and time \cite{costello2009bacterial}.
Inter-individual versus intra-individual variation \cite{hmp2012structure}.

When we perturb the microbiome, sometimes it makes fundamental changes.
Antibiotics fundamentally disturb microbial communities, resulting in an effect that lasts years.
\cite{jernberg2007long,dethlefsen2008pervasive,jakobsson2010short,dethlefsen2011incomplete}

Sometimes these changes lead to phenotypic changes in the host.
There is strong evidence of interaction between the gut microbiome and obesity, but the story is yet to be fully explored.
An intervention study has established human gut microbes associated with obesity \cite{ley2006microbial}.
An obese phenotype can be transferred from mouse to mouse by gut microbiome transplantation \cite{turnbaugh2006obesity}.
Pregnant human gut microbiome \cite{koren2012host} leads to metabolic disorder in mice.

These findings have yet to find their way into a real understanding.
Amish study \cite{zupancic2012analysis}
"Neither BMI nor any metabolic syndrome trait was associated with a particular gut community. Network analysis identified twenty-two bacterial species and four OTUs that were either positively or inversely correlated with metabolic syndrome traits, suggesting that certain members of the gut microbiota may play a role in these metabolic derangements."

In order to really get an understanding, researchers are shifting from community dynamics to fine-scale interactions between host and microbe \cite{hooper2012interactions}.
Some major findings there.

So far we have been focusing on bacteria, though archaea also play a role.
Archaea have also been found in the gut.

Although the focus of much of the work has been on bacteria and archaea, other groups of life have attracted more attention lately, such as the fungi.

Viruses too
febrile study \cite{wylie2012sequence}

@@

@@

\section{Discussion}
What we can expect next.
As described above, we can expect much more development in terms of specifics.
\cite{hooper2012interactions}
Except to the extent that it is useful to look at gene trees to understand function, phylogenenetics doesn't really have a role.

But in addition to deeper probing we can also expect wider coverage.
Lots of data.
Citizen microbiome

Comparative studies of microbiomes.
\cite{phillips2012microbiome}
Microbiome analysis among bats describes influences of host phylogeny, life history, physiology and geography

Clinical applications,
\cite{clarridge2004}
{{I}mpact of 16{S} r{RNA} gene sequence analysis for identification of bacteria on clinical microbiology and infectious diseases}
Understand cause and effect of diseases.
Take an ecololgical view on atopic dermititis.

For routine identification of a single microbe in culture, MALDI-TOF kicks ass and is cheap.

The field has experienced a frenetic rate of expansion over the past xxx years, and sometimes it can feel like researchers in this area have forgotten some of the basic principles of study design, such as power calculations, that are well recognized in other areas of medical research.
Although the human microbiome has led to a lot of "hype," the microbes inside of us and research on them is here to stay.
Thus we can look forward to the field of human microbiome analysis settling down to a comfortable and mature middle age as an interesting intersection between ecology and medicine.

### 2013-06-05 trash.tex
cranking away...

\subsection{Software pipelines}
\cite{caporaso2010qiime}
\cite{schloss2009introducing}
\cite{lynch2013axiome}

>>>>

Substantial attention was given to whether microbial ``types'' exist.
A high-profile paper from Per Bork's group had enterotypes: Bacteroides, Prevotella or Ruminococcus.
\cite{arumugam2011enterotypes}
Appears that there is a real tendency to cluster around Bacteroides or Prevotella \cite{wu2011linking}
